---
title: "Netflix Data Analysis"
subtitle: "From Data Cleaning to Advanced Insights"
author: "Petra Dudas and Flaminia Neda"
format:
  revealjs:
    theme: dark
    slide-number: true
    transition: fade
    css: styles.css
execute:
  echo: false
  warning: false
  message: false
editor: visual
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(lubridate)

netflix_raw <- read_csv("data/netflix_titles.csv")

netflix_clean <- netflix_raw %>%
  mutate(
    director = if_else(is.na(director), "Unknown", director),
    cast = if_else(is.na(cast), "Unknown", cast),
    country = if_else(is.na(country), "Unknown", country),
    rating = if_else(is.na(rating), "Unknown", rating)
  ) %>%
  filter(!is.na(date_added), !is.na(duration)) %>%
  mutate(
    date_added = mdy(date_added),
    duration_number = as.numeric(str_extract(duration, "\\d+")),
    duration_type = case_when(
      str_detect(duration, "Season") ~ "Seasons",
      str_detect(duration, "min") ~ "Minutes",
      TRUE ~ "Unknown"
    )
  )
```

## Introduction

-   Netflix is one of the world's leading streaming platforms, offering a wide variety of movies and TV shows across different genres and countries.

-   In this project, we analyze a Netflix dataset to uncover patterns in content distribution, genres, and trends over time.

## Data Overview

The dataset contains information about Netflix titles, including:

‚ÄĘ‚Ā† ‚Ā†Title type (Movie or TV Show)

‚ÄĘ‚Ā† ‚Ā†Country of production

‚ÄĘ‚Ā† ‚Ā†Release year

‚ÄĘ‚Ā† ‚Ā†Genre classification

‚ÄĘ‚Ā† ‚Ā†Duration

## Data Cleaning Process

Before analyzing the data, several cleaning steps were required:

‚ÄĘ‚Ā† ‚Ā†Splitting duration into numeric and categorical values

‚ÄĘ‚Ā† ‚Ā†Handling missing values

‚ÄĘ‚Ā† ‚Ā†Separating multiple countries into individual entries

‚ÄĘ‚Ā† ‚Ā†Removing duplicates

## Top Producing Countries

```{r}
library(dplyr)
library(ggplot2)
library(gganimate)
library(gifski)

top5_countries <- netflix_clean %>%
  filter(!is.na(country),
         country != "Unknown",
         country != "") %>%
  count(country, sort = TRUE) %>%
  slice_head(n = 5) %>%
  pull(country)

country_year <- netflix_clean %>%
  filter(!is.na(country),
         country != "Unknown",
         country != "",
         country %in% top5_countries,
         !is.na(release_year),
         release_year >= 2000) %>%
  count(release_year, country)

ggplot(country_year,
       aes(x = reorder(country, n), y = n)) +
  geom_col(fill = "#E50914") +
  coord_flip() +
  labs(
    title = "Top 5 Countries on Netflix",
    subtitle = "Release year: {closest_state}",
    x = "Country",
    y = "Number of Titles"
  ) +
  theme_minimal() +
  transition_states(
    release_year,
    transition_length = 2,
    state_length = 1
  ) +
  ease_aes("cubic-in-out")

```

The United States clearly dominates Netflix content production, followed by India and the United Kingdom. This highlights the strong influence of these countries in the global streaming market.

## Most Popular Genres

```{r}
netflix_clean %>%
  count(listed_in, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = reorder(listed_in, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top Genres on Netflix",
       x = "Genre",
       y = "Count")
```

International Movies are the most common genre on Netflix, followed by Drama and Comedy. This suggests a strong focus on globally diverse content.

## Growth of Netflix Over Time

```{r}
titles_per_year <- netflix_clean %>%
  filter(!is.na(release_year)) %>%
  count(release_year)

ggplot(titles_per_year, aes(x = release_year, y = n)) +
  geom_point(color = "#E50914", size = 2) +
  geom_smooth(method = "lm", se = TRUE, color = "#E50914") +
  labs(
    title = "Growth of Netflix Content Over Time",
    subtitle = "Linear regression showing the trend in released titles",
    x = "Release Year",
    y = "Number of Titles"
  ) +
  theme_minimal()

```

The regression reveals a strong growth pattern, which forms the basis for the future predictions shown next.

## Future Prediction

```{r}
model_titles <- lm(n ~ release_year, data = titles_per_year)

future_years <- data.frame(
  release_year = c(2022, 2023, 2024, 2025)
)

prediction_table <- future_years %>%
  mutate(
    predicted_titles = round(predict(model_titles, newdata = future_years), 0)
  )

prediction_table %>%
  knitr::kable(
    col.names = c("Year", "Predicted Titles"),
    align = "cc"
  ) %>%
  kableExtra::kable_styling(
    full_width = FALSE,
    position = "center",
    font_size = 40
  )

```

## Content by Age Group

```{r}
netflix_clean <- netflix_clean %>%
  mutate(age_group = case_when(
    rating %in% c("TV-Y", "TV-Y7", "G") ~ "Kids",
    rating %in% c("PG", "TV-PG") ~ "Family",
    rating %in% c("PG-13", "TV-14") ~ "Teens",
    rating %in% c("R", "TV-MA", "NC-17") ~ "Adults",
    TRUE ~ "Other"
  ))
```

```{r}
netflix_clean %>%
  count(age_group) %>%
  ggplot(aes(x = age_group, y = n, fill = age_group)) +
  geom_col() +
  labs(
    title = "Netflix Content by Age Group",
    x = "Age Group",
    y = "Number of Titles"
  ) +
  theme(legend.position = "none")
```

Most Netflix content is targeted toward adults and teenagers, while significantly less content is designed specifically for children. This reflects Netflix‚Äôs primary audience focus.

## Movie Recommendations by Age Group

```{r}
library(dplyr)
library(knitr)
library(kableExtra)

recommendations <- netflix_clean %>%
  filter(type == "Movie") %>%
  group_by(age_group) %>%
  slice_head(n = 3) %>%
  select(Age_Group = age_group, Title = title, Rating = rating)

kable(recommendations) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    font_size = 18
  )
```

## Relationship Between Release Year and Duration

```{r}
library(dplyr)
library(ggplot2)
library(stringr)

netflix_duration <- netflix_clean %>%
  filter(type == "Movie") %>%
  mutate(duration_num = as.numeric(str_extract(duration, "\\d+"))) %>%
  drop_na(duration_num)

netflix_duration %>%
  group_by(release_year) %>%
  summarise(avg_duration = mean(duration_num), .groups = "drop") %>%
  ggplot(aes(x = release_year, y = avg_duration)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.5) +
  labs(
    title = "Average Movie Duration Over Time",
    x = "Release Year",
    y = "Average Duration (minutes)"
  ) +
  theme_minimal()
```

## Conclusion

-   The analysis highlights key patterns in Netflix‚Äôs content strategy, including the dominance of certain countries, the popularity of specific genres, and the platform‚Äôs rapid growth in recent years.

-   While Netflix continues to expand its content library, core characteristics such as movie duration remain consistent, reflecting stable industry standards.

-   Overall, the findings demonstrate how data analysis can be used to uncover meaningful insights and support content-related decision-making.

## Key Takeaways

‚ÄĘ‚Ā† ‚Ā†Netflix content is heavily concentrated in a few major production countries

‚ÄĘ‚Ā† ‚Ā†The platform primarily targets adult and teenage audiences

‚ÄĘ‚Ā† ‚Ā†Content diversity is reflected in the popularity of international movies

‚ÄĘ‚Ā† ‚Ā†Data-driven approaches can support recommendation systems and strategic decisions

‚ÄĘ‚Ā† ‚Ā†Using Quarto :)
